---
title: "Running a simulation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Running a simulation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

This vignette will show you how to create a single simulation using the basic
functions from the package. To create a DCM simulation we need: a test, a 
Q-matrix and examinees. We can do all these quickly using saltr. 

## Creating our test

Let's first create a Q-matrix.

```{r}
library(saltr)

qmatrix <- create_qmatrix(num_attr=3, items_per_type = c(2,1))
qmatrix
```

The `create_qmatrix` objective is give us a fast template for a Q-matrix which 
we then can modify. Let's make item 9 measure all attributes.

```{r}
qmatrix[9,] <- c(1,1,1) 
```

Perfect. So far, so good. Now, we want to create a test that has to be compatible
with our Q-matrix. This is normally a tricky thing to do, but using `build_test_parameters` the operation is simplified.

```{r}
# We pass our qmatrix as an argument 
test <- build_test_parameters(qmatrix)
test

# Item 8 has the intercept 0, the main effects 1 and 3, 
# and the two-way interaction 1-3
test[8,]
```

Because we did not pass any extra arguments, the result of `build_test_parameters` 
is just a mask of the parameters that should be present in the test. The columns 
represent the attribute parameters. We can also pass the values we want for each
interaction level starting with the intercept.

```{r}
# Only intercept and main effects
test <- build_test_parameters(qmatrix, -2, 2)
test

# Intercept, main effects and two-way interactions
test <- build_test_parameters(qmatrix, -2, 2, .5)
test
```

In the same way that before, this function greatly simplifies the creation of a
starting template for the test which can be then adjusted. Notice that the 3-way 
interaction effect of item 9 is 0 and not NA, this is important for compatibility 
with the matrix. For example, let's make one of the main effects of items 6 and 7 smaller.

```{r}
test[6, 4] <- 1
test[7, 3] <- 1

# Using pretty_print makes it easier to read
pretty_print(test)
```

## Creating our examinees

Great! now we have our test we need our examinees. This is a function that has a 
random component so we use a `generate_*` function. This function is a wrapper for
the `bindata::rmvbin` with some extra logic for errors and conciseness. For example,
if only one value is passed, it assumes they are all the same.  

```{r}
examinees <- generate_examinees(
  1000, # sample size
  total_attrs = 3,
  base_rate = .5, # All three attributes will have the same base-rate
  attr_corr = c(.3, 0, 0)
)

# Checking results
colMeans(examinees)
cor(examinees) |> round(2)
```

An important attribute is `strict`, if it is `FALSE`, the attribute 
correlation is assumed to be  the correlation of the underlying normal variables. 
By default this value is  `TRUE`, meaning it will try and create a binary matrix 
that follows those restrictions.If not possible it will throw an error.
 
```{r, , warning=TRUE, error=TRUE}
# This will throw an error
sample_1 <- generate_examinees(
  1000, 
  total_attrs = 3,
  base_rate = c(.8, .5, .2),
  attr_corr = c(.8, .1, 0),
  strict = TRUE
)

# This wont
sample_2 <- generate_examinees(
  1000, 
  total_attrs = 3,
  base_rate = c(.8, .5, .2),
  attr_corr = c(.8, .1, 0),
  strict = FALSE
)

# Although it is always better to pass some value for tolerance
# To get info on the generation.
sample_3 <- generate_examinees(
  1000, 
  total_attrs = 3,
  base_rate = c(.8, .5, .2),
  attr_corr = c(.8, .1, 0),
  tolerance = .1
)
```

## Responses

Now that we have a test and our examinees we can generate our responses. We do
not need to pass a Q-matrix as that information is already encoded in the test.
That way there's less chances of making mistakes. 

```{r}
responses <- generate_responses(examinees, test)
```

A nice feature of this function is that it can help us avoid mistakes that could
have happened during the modification of the test. For example, what if we 
inadvertently added a second main effect on item 2.

```{r, error=TRUE}
wrong_test <- test

# Now item 2 has a main effect but the interaction 1-2 is NA 
wrong_test[2,3] <- .5

responses <- generate_responses(examinees, wrong_test)
```

And now we can just fit our model.

## Wrapping it all together

Our complete code to start from zero and get simulated responses to a test is 
just a couple of lines!

```{r}
qmatrix <- create_qmatrix(num_attr=3, items_per_type = c(2,1))
qmatrix[9,] <- c(1,1,1) 

test <- build_test_parameters(qmatrix, -2, 2, .5)
test[6, 4] <- 1
test[7, 3] <- 1

examinees <- generate_examinees(
  1000, 
  total_attrs = 3,
  base_rate = .5, 
  attr_corr = c(.3, 0, 0)
)

responses <- generate_responses(examinees, test)
```


