---
title: "Running multiple simulations"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Running multiple simulations}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message=FALSE, include=FALSE}
library(saltr)
library(tidyverse)
library(CDM)
```


Let's create a simulation to measure the precision of the parameter estimation for DINO items when we don't have simple items.

## Setting the base design

First we create a Q-matrix and a sample of respondents. 

```{r}
library(saltr)

qmatrix <- create_qmatrix(3, c(0,2))
sample <- generate_sample(1000, 3, base_rate = .5)
```

We can create all items to be the same for simplicity.

```{r}
item_probs <- matrix(
  rep(c(.2, .7, .7, .7), 6), # DINO item, non-master .2 probability, masters .7
  nrow=6, byrow=TRUE
)

test <- create_test(item_probs)
test
```

Let's simulate one set of responses: 

```{r}
responses <- generate_responses(qmatrix, sample, test)
head(responses)
```

Now, because this a simulation we need to: generate multiple responses, fit a model and somehow save the results.

## Creating a simulation function

We can use a simple function like this. This function encapsulates the creation of both the responses and the fitting of the model. This is key for using `doFuture`.

```{r}
simulation <- function() {
  
  # Generating a new set of responses
  responses <- generate_responses(qmatrix, sample, test)

  # Fitting the model 
  model <- CDM::gdina(
    responses, 
    qmatrix, 
    linkfct = "logit", 
    progress = FALSE)

}
```

We can then create a function which will return a simple value that we can save and report on.

```{r}
extract_parameters <- function(model) {
  parameters <- model$coef[ , c(3,4,6)] 
  intercepts <- parameters[parameters$partype==0, 3]
  main_effects <- parameters[parameters$partype %in% c(1,2), 3]
  
  return(
    c(mean(intercepts),
      mean(main_effects))
  )
}
```

The new simulation function is now this:

```{r}
simulation <- function() {
  
  # Generating a new set of responses
  responses <- generate_responses(qmatrix, sample, test)

  # Fitting the model 
  # We supress the warnings of the simulations with all 1 or 0s.
  model <-suppressWarnings(CDM::gdina(
    responses, 
    qmatrix, 
    linkfct = "logit", 
    progress = FALSE))

  # Now we return a measure, not the model itself.
  results <- extract_parameters(model)
  return(results)
}
```

We can run it to see it's working as expected:
```{r, eval=FALSE}
simulation()
```

## Iterating

We can now use this function to iterate and save all the results, for example like this:

```{r eval=FALSE}
reps <- 10

results <- matrix(NA, ncol=2, nrow=reps)

for (i in 1:reps) {
  results[i, ] <- simulation()  
}
```

## Parallel processing

saltr can be used together with the [doFuture](https://cran.r-project.org/web/packages/doFuture/vignettes/doFuture-1-overview.html) package to take advantage of parallel processing through the library `doFuture`with very few changes.

Although, it comes with some overhead so it's benefits don't really show until 
we do bigger simulations.

```{r eval=FALSE}
library(doFuture)
reps <- 10

results <- foreach(
  rep = 1:reps, .combine = rbind, .options.future = list(seed = TRUE)
  ) %dofuture% {
    simulation()
}
```


